COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region

### 2020-03-23 draft_analysis_Veneto_html.Rmd
Add home page, navigation and theme

```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
---
title: "COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output:html_document:
  df_print: paged
pdf_document: 
  toc: yes
---
<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

@@

```{r}
p <- dygraph(dat_csv_dy,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800) 
p
```

### The S(E)IR model (to be revised)

@@

$$
\widehat\beta_1=(\widehat{R_0}-1)\,(\gamma+\mu)
$$
The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
$$
\hat{R_0}=1+\frac{\hat{\beta_1}}{\gamma}
$$
Respect to the SIR model, $R_0$ can be estimated as follows:
$$
\hat{R_0}=\frac{\hat{\beta}}{\gamma}
$$
And this was we can retrive the value of $\beta$ in the SEIR model by means of 
$$
\hat{R_0}=\frac{\hat{\beta}}{\gamma}=1+\frac{\hat{\beta_1}}{\gamma}\\
\hat{\beta}=(1+\frac{\hat{\beta_1}}{\gamma})*{\gamma}={\gamma}+\hat{\beta}_1
$$
where $\beta_1$ is the slope coefficient. \\

@@

However, in the calculation of R0 we considered an infectious period of 21 days for two reasons:  
1) the majority of cases is asymptomatic;
2) 21 days is the worst scenario because in this period COVID19 symptoms can be confused with the concomitant FLU epidemic (https://www.webmd.com/lung/news/20200310/know-the-symptoms-of-covid19).

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
```{r R0 forecast}

time<-3:(days-2)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian)
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
predict <- data.frame(beta_vec=c(beta_vec,rep(NA,forecast+4)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
beta.predict<-predict 
r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*21+1
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

@@

```{r ini parameters plot, echo=TRUE}
# initial number of infectus
I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
# initial number of recovered, based on the proportion of discharged from the health services
prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
# Veneto population
N=4905864
# duration of COVID19 as the sum of incubation (5-14 days) and the duration of diseases (about 20 days)
duration<-21
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```

@@

```{r estimation of gamma and exposed, echo=TRUE}
# number of the last considered days for calibration
last<-5
#I_start
I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
#R_start
R_1<-prop*dat_csv$totale_casi[(days-last)]
#loglikelihood function
LogLikfun <- function (initials,parameters,obs) {
n<-length(obs)
N=4905864
seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
#Poisson
#sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
#Neg Binomial
sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
#SSE
#-sum((obs[-1]-seir_fit$results$I[-1]*N)^2)
}
### logit and its inverse for sigma 
logit <- function (p) log(p/(1-p))    # the logit transform
expit <- function (x) 1/(1+exp(-x))   # inverse logit

f1<-function(par){
f_exp<-expit(par[1])
sigma<-expit(par[2])
parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma =1/duration)
initials <- c(S = (1-(f_exp+I_1/N-R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
-LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
}
est<-optim(fn=f1,par=c(logit(0.0005),logit(0.1)))
#estimated par
expit(est$par)
f_exp<-expit(est$par[1])
sigma<-expit(est$par[2])
parameters <- c(mu = mu0, beta =(mean(beta_vec[1:(days-last-4)])+1/duration), sigma = sigma, gamma = 1/duration)
initials <- c(S = (1-(f_exp+I_1/N-R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
pro<-SEIR(pars = parameters, init = initials, time = 0:last)
#predicted vs observed
cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
```

@@

```{r status, echo=TRUE}
#mthe unknowrn fraction of exposed people is
f_exp;
#that is 
f_exp/(I_1/N)
#times the infected
```
For the beta parameter, we perfom a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
We estimate a BSTS model specifing a local linear trend ans 1000 simulation.

@@

```{r bsts model fit and prediction, echo=TRUE}
# Bayesian Structural Time Series
ss <- AddLocalLinearTrend(list(), beta_vec)
model1 <- bsts(beta_vec,
               state.specification = ss,
               niter = 1000)
par(mfrow = c(1,1))
plot(model1, "components", ylab="Beta1 coefficient", xlab="days")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions")
# matrix of beta coefficients
coef<-(pred1$distribution[,3:16])
par(mfrow = c(1,1))
```
For each vector of simulated beta coefficients we perform a SEIR model.
We save the results and plot a credible interval at 50% (25-75% percentile).

```{r  scenario plot , echo=TRUE}
seir1_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir1<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
    if( i==1) initials <- c(S = 1-(f_exp+I_start/N+R_start/N), E = f_exp, I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
    seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))

@@

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Veneto Region: Scenario  (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

@@

The estimated (median) numbers of the current scenario by date are:
```{r  scenario numbers, echo=TRUE}
S_seir_med<-tapply(seir1_sim$S,seir1_sim$time,median)
E_seir_med<-tapply(seir1_sim$E,seir1_sim$time,median)
R_seir_med<-tapply(seir1_sim$R,seir1_sim$time,median)
forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
rownames(forecast)<-days.ahead
kable(forecast)
```

>>>>

```{r echo=FALSE}
library(knitr)
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_knit$set(root.dir= normalizePath('..'))
knitr::opts_chunk$set(error = FALSE)
```
---
title: "COVID19 - Forecast and predictions using a time dependent SEIR model for the Veneto Region"
author: "Paolo Girardi"
date: "`r format(Sys.time(), '%d %B, %Y')`"
output: rmarkdown::github_document
always_allow_html: yes
---
<a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/"><img alt="Creative Commons License" style="border-width:0" src="https://i.creativecommons.org/l/by-nc/4.0/88x31.png" /></a><br />This work is licensed under a <a rel="license" href="http://creativecommons.org/licenses/by-nc/4.0/">Creative Commons Attribution-NonCommercial 4.0 International License</a>.

@@

```{r}
p <- dygraph(dat_csv_dy,main=paste("Veneto Region",sep =""),xlab="Day",height=400,width=800) 
p
```
The time series of COVID-19 in the Veneto Region was influenced by the presence of a hotspot in the small village of Vo Euganeo.

```{r}
plot(days_dy[-1],diff(dat_csv$totale_casi), ylab="New cases",xlab="Date",type="l")
```
### The S(E)IR model (to be revised)

@@

$$
\widehat\beta_1=(\widehat{R_0}-1)\,(\gamma+\mu)
$$
The parameter $\mu$<<$\gamma$ and it can not be considered. As consequence, R0 can be estimated as follows
$$
\hat{R_0}=1+\frac{\hat{\beta_1}}{\gamma}
$$
Respect to the SIR model, $R_0$ can be estimated as follows:
$$
\hat{R_0}=\frac{\beta}{\gamma}
$$
And this was we can retrive the value of $\beta$ in the SEIR model by means of 
$$
\hat{R_0}=\frac{\hat{\beta}}{\gamma}=1+\frac{\hat{\beta_1}}{\gamma}\\
\hat{\beta}=(1+\frac{\hat{\beta_1}}{\gamma})*{\gamma}={\gamma}+\hat{\beta}_1
$$
where $\beta_1$ is the slope coefficient. \\

@@

However, in the calculation of R0 we considered an infectious period of 18 days on the basis of recent publications (https://www.nature.com/articles/s41421-020-0148-0).

@@

The R0 shows a decreasing trend in the last period. We use the estimated trend between R0 and time to calculate the future R0 value for the next 14 days. 
We predict beta (and R0) for the next 14 days by means of a linear regressione model, assuming a Normal distribution for the beta (the slope).
```{r R0 forecast}
#start from the day 7, since vo hotspot can influence the trend
time<-3:(days-2)
weekend<-rep(c(1,2,3,4,5,6,7),ceiling(days/7))[3:(days-2)]
data<-data.frame(time)
beta.model<-glm(beta_vec~time,weights = 1/sd_vec,family=gaussian,data)
summary(beta.model)
anova(beta.model,test="Chisq")
#there is an effect of the week, as supposed 
#we can use
#summary(gamlss(formula = beta_vec ~ time + bs(weekend, 3),      weights = 1/sd_vec,Lp=1) )
forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(beta.model,type='response',newdata=data.frame(time=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
predict <- data.frame(beta_vec=c(rep(NA,2),beta_vec,rep(NA,forecast+2)),time=date,fit=pre$fit,lwr=pre$fit-1*1.96*pre$se.fit,upr=pre$fit+1*1.96*pre$se.fit)
beta.predict<-predict 
r0.predict<-beta.predict
r0.predict[,c(1,3:5)]<-r0.predict[,c(1,3:5)]*duration+1
# plot the points (actual observations), regression line, and confidence interval
p <- ggplot(r0.predict, aes(date,beta_vec))
p <- p + geom_point() +labs(x="Date",y="R0 value") 
p <- p + geom_line(aes(date,fit))
p <- p + geom_ribbon(aes(ymin=lwr,ymax=upr), alpha=0.3)
p
```

@@

```{r ini parameters plot, echo=TRUE}
# initial number of infectus
I_start<-dat_csv$totale_attualmente_positivi[dim(dat_csv)[1]]; I_start
# initial number of recovered, based on the proportion of discharged from the health services
prop<-dat_csv$dimessi_guariti[dim(dat_csv)[1]]/dat_csv$totale_ospedalizzati[dim(dat_csv)[1]]
R_start<-prop*dat_csv$totale_casi[dim(dat_csv)[1]]; R_start
# Veneto population
N=4905864
#  duration of COVID19 diseases 
duration<-18
#mortality rate 
mu0<-1/(82*365.25) # 1/lifespan
```

@@

```{r estimation of gamma and exposed, echo=TRUE}
# number of the last considered days for calibration
last<-10
#I_start
I_1<-dat_csv$totale_attualmente_positivi[(days-last)]
#R_start
R_1<-prop*dat_csv$totale_casi[(days-last)]
#loglikelihood function
LogLikfun <- function (initials,parameters,obs) {
n<-length(obs)
N=4905864
seir_fit <- SEIR(pars = parameters, init = initials, time = 0:last)
#Poisson
sum(dpois(x=obs[-1],lambda = seir_fit$results$I[-1]*N,log=TRUE))
#Neg Binomial
#sum(dnbinom(x=obs[-1],mu=seir_fit$results$I[-1]*N,size=N,log=TRUE))
#SSE
#-sum((obs[-1]-seir_fit$results$I[-1]*N)^2)
}
### logit and its inverse for sigma 
logit <- function (p) log(p/(1-p))    # the logit transform
expit <- function (x) 1/(1+exp(-x))   # inverse logit

f1<-function(par){
f_exp<-expit(par[1])
#max 0.33 according to https://www.nejm.org/doi/10.1056/NEJMoa2001316 is 1/5.2
sigma<-expit(par[2])/3
parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma =1/duration)
initials <- c(S = (1-(f_exp+I_1/N+R_start/N)), E = f_exp, I = I_1/N, R = R_1/N)
-LogLikfun(initials,parameters,dat_csv$totale_attualmente_positivi[c((days-last):days)])
}
est<-optim(fn=f1,par=c(logit(0.02),logit(0.2)))
#estimated par
expit(est$par)

f_exp<-expit(est$par[1])
sigma<-expit(est$par[2])/3
parameters <- c(mu = mu0, beta =(as.numeric(beta_vec[days-last-2])+1/duration), sigma = sigma, gamma = 1/duration)
initials <- c(S = (1-(f_exp+I_1/N+R_1/N)), E = f_exp, I = I_1/N, R = R_1/N)
pro<-SEIR(pars = parameters, init = initials, time = 0:last)
#predicted vs observed
cbind(pro$results$I*N,dat_csv$totale_attualmente_positivi[c((days-last):days)])
```

@@

```{r status, echo=TRUE}
#mthe unknowrn fraction of exposed people is
pro$results$E[last+1]
#that is 
pro$results$E[last+1]/(I_1/N)
#times the infected
```

# Forecast 1 - Results based on beta coefficient trend

For the beta parameter, we perfomed a simulation on its trend by means of a Bayesian Structural Time Series using the library bsts of R.    
We estimated a BSTS model specifing a local linear trend and a seasonality component. The model predicted 1000 trajectories forward of the beta coefficent.

@@

```{r bsts model fit and prediction, echo=TRUE}
# Bayesian Structural Time Series
#we drop the first 6 beta_vec
ss <- AddLocalLinearTrend(list(), beta_vec[6:(days-4)])
ss <- AddSeasonal(ss, beta_vec[6:(days-4)], nseasons = 7)
model1 <- bsts(beta_vec[6:(days-4)],
               state.specification = ss,
               niter = 1000,seed=123)
par(mfrow = c(1,1))
plot(model1, "components")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
# matrix of beta coefficients
coef<-(pred1$distribution[,3:16])
par(mfrow = c(1,1))
```

For each vector of simulated beta coefficients we perform a SEIR model.
We save the results and plot a credible interval at 50% (25-75% percentile).

```{r  scenario plot , echo=TRUE}
seir1_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir1<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir1_temp$results$S[2], E = seir1_temp$results$E[2], I =seir1_temp$results$I[2], R = seir1_temp$results$R[2])
    seir1_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir1 <- rbind(seir1,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir1_sim<-rbind(seir1_sim,cbind(rep(s,forecast),seir1))

@@

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

@@

The estimated (median) numbers of the current scenario by date are:

```{r  scenario numbers, echo=TRUE}
S_seir_med<-tapply(seir1_sim$S,seir1_sim$time,median)
E_seir_med<-tapply(seir1_sim$E,seir1_sim$time,median)
R_seir_med<-tapply(seir1_sim$R,seir1_sim$time,median)
forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
rownames(forecast)<-days.ahead
kable(forecast)
```

# Scenario 2 - Forecast based on the last beta coefficient

We estimate a second scenario using a Random Walk on the basis of the last value of the beta.

```{r  scenario2 sim and plot , echo=TRUE}
ss<-NULL
set.seed(123)
ss <- AddLocalLinearTrend(list(), rnorm(100,beta_vec[days-4],sd_vec[days-4]))
model1 <- bsts(beta_vec[6:(days-4)],
               state.specification = ss,
               niter = 1000,seed=123)
par(mfrow = c(1,1))
plot(model1, "components")
#previsioni
pred1 <- predict(model1, horizon = 16, burn = 100)
par(mfrow = c(1,1))
plot(pred1 , ylab="Beta1 coefficient",main="Data and predictions",ylim=c(-1,1)*IQR(pred1$distribution)+median(pred1$distribution))
# matrix of beta coefficients
coef<-(pred1$distribution[,3:16])
par(mfrow = c(1,1))

seir2_sim<-NULL
for(s in 1:dim(coef)[1]){
  # average number of single connections of an infected person
  # less contacts, less probability of new infections
  # we keep constant the other parameters
  forecast<-14
  seir2<-NULL
  for(i in 1:forecast){
    parameters <- c(mu = mu0, beta = (matrix(coef[s,i])+1/duration), sigma = sigma, gamma = 1/duration)
    if( i==1) initials <- c(S = 1-(pro$results$E[last+1]+I_start/N+R_start/N), E = pro$results$E[last+1], I = I_start/N, R = R_start/N)
    if( i>1) initials <- c(S = seir2_temp$results$S[2], E = seir2_temp$results$E[2], I =seir2_temp$results$I[2], R = seir2_temp$results$R[2])
    seir2_temp <- SEIR(pars = parameters, init = initials, time = 0:1)
    seir2 <- rbind(seir2,SEIR(pars = parameters, init = initials, time = 0:1)$results[2,])
  }
  seir2_sim<-rbind(seir2_sim,cbind(rep(s,forecast),seir2))

}
seir2_sim[,2]<-rep(1:forecast,dim(coef)[1])
colnames(seir2_sim)[1]<-"sim"

### confidence limits
I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1
mu.lower<-c(dat_csv$totale_attualmente_positivi,I_seir_lwr*N)
mu.upper<-c(dat_csv$totale_attualmente_positivi,I_seir_upr*N)
mu.med<-xts(c(dat_csv$totale_attualmente_positivi,I_seir_med*N),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Credible Interval ",100*0.50,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
```

At the end of the 2 weeks (`r date[length(date)]`):
*the number of infected is  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N)[forecast]`).

*the total number of COVID19 cases is expected to be  (`r (tapply(seir2_sim$I,seir2_sim$time,median)*N+tapply(seir2_sim$R,seir2_sim$time,median)*N)[forecast]`).

The estimated (median) numbers of the second scenario by date are:
```{r  scenario2 numbers, echo=TRUE}
S_seir_med<-tapply(seir2_sim$S,seir2_sim$time,median)
E_seir_med<-tapply(seir2_sim$E,seir2_sim$time,median)
R_seir_med<-tapply(seir2_sim$R,seir2_sim$time,median)
forecast<-data.frame(S_seir_med,E_seir_med,I_seir_med,R_seir_med)*N
colnames(forecast)<-c("Susceptible","Exposed","Infected","Removed")
rownames(forecast)<-days.ahead
kable(forecast)
```

# Number of admitted in intensive care

The number of person admitted in intensive care repart is a fraction of the COVID 19 case.
The trend can be estimated by means a binomial regression model.  
We report the number of intensive care patients on the basis of the estimated COVID19 cases of  the previous scenario 1 and 2.

```{r  intensive care numbers, echo=TRUE}
fit_int_care<-glm(cbind(terapia_intensiva,totale_attualmente_positivi-terapia_intensiva)~t,family=binomial,data=dat_csv)

forecast=14
# add 'fit', 'lwr', and 'upr' columns to dataframe (generated by predict)
pre<-predict(fit_int_care,type='response',newdata=data.frame(t=1:(days+forecast)),se.fit=TRUE)
date<-seq(as.Date("2020-02-24"),as.Date("2020-02-24")+forecast-1+dim(dat_csv)[1],1)
plot(date,pre$fit,type="l",main="Proportion of admitted COVID 19 in intensive care",ylim=range(dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi))
points(date[1:days],dat_csv$terapia_intensiva/dat_csv$totale_attualmente_positivi)
lines(date,pre$fit+1.96*pre$se.fit,col=3,lty=2)
lines(date,pre$fit-1.96*pre$se.fit,col=3,lty=2)
legend("topright",c("fitted","95% Confidence Interval"),lty=1:2,col=c(1,3))

days.before<-date[1:days]
days.ahead<-date[(days+1):(days+forecast)]
step.ahead<-forecast+1

#Scenario 1

I_seir_med<-tapply(seir1_sim$I,seir1_sim$time,median)
I_seir_lwr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.25)
I_seir_upr<-tapply(seir1_sim$I,seir1_sim$time,quantile,p=0.75)

mu.lower<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit-1.645*pre$se.fit)[(days+1):(days+forecast)]))
mu.upper<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit+1.645*pre$se.fit)[(days+1):(days+forecast)]))
mu.med<-xts(c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit)[(days+1):(days+forecast)])),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Veneto Region: Scenario 1 (Confidence Interval ",100*0.90,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
#Numbers
intensive_care<-data.frame(mu.med,mu.lower,mu.upper)[(days+1):(days+forecast),]
kable(intensive_care)

#Scenario 2

I_seir_med<-tapply(seir2_sim$I,seir2_sim$time,median)
I_seir_lwr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.25)
I_seir_upr<-tapply(seir2_sim$I,seir2_sim$time,quantile,p=0.75)

mu.lower<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit-1.645*pre$se.fit)[(days+1):(days+forecast)]))
mu.upper<-c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit+1.645*pre$se.fit)[(days+1):(days+forecast)]))
mu.med<-xts(c(dat_csv$terapia_intensiva,I_seir_med*N*((pre$fit)[(days+1):(days+forecast)])),order.by = c(days.before,days.ahead),frequency = 7)
counts<-mu.med
mu<-xts(x = as.matrix(cbind(counts,mu.lower,mu.upper)) , order.by = c(days.before,days.ahead))
p <- dygraph(mu,main=paste("Veneto Region: Scenario 2 (Confidence Interval ",100*0.9,"%)",sep = ""),ylab=" Infected",xlab="Day",height=400,width=800) %>%  dySeries(c("mu.lower", "counts", "mu.upper"),label="counts")
p<-p %>% dyLegend(show = "always", hideOnMouseOut = FALSE) %>%  dyShading(from = days.ahead[1], to = days.ahead[step.ahead], color = "#CCEBD6")%>% dyEvent(days.ahead[1], "Prediction", labelLoc = "bottom")
p
#Numbers
intensive_care<-data.frame(mu.med,mu.lower,mu.upper)[(days+1):(days+forecast),]
kable(intensive_care)
```
